---
title: "An Attempt At The Titanic Competition"
output:
  html_document:
    toc: yes
  html_notebook:
    number_sections: yes
    toc: yes
---
# Introduction
This notebook is meant to serve as a playground for me to play with the `h2o` package.

## Libraries and data
Load the required libraries:
```{r}
library(magrittr)
library(ggplot2)
library(scales)
# library(caret)
library(h2o)
```
Import the data set:
```{r}
train.df <-
  read.csv(
  './input/train.csv',
  colClasses = c(
  'character',
  'factor',
  'factor',
  'character',
  'factor',
  'numeric',
  'integer',
  'integer',
  'character',
  'numeric',
  'factor',
  'factor'
  ),
  na.strings = c('')
  )

test.df <-
  read.csv(
  './input/test.csv',
  colClasses = c(
  'character',
  'factor',
  'character',
  'factor',
  'numeric',
  'integer',
  'integer',
  'character',
  'numeric',
  'factor',
  'factor'
  )
  )
```

# Data Pre-Processing
## Feature Engineering
Combine the train and test data frames:
```{r}
consol.df <- (train.df %>%
                dplyr::mutate(label = 'train')) %>%
                dplyr::bind_rows((test.df %>%
                dplyr::mutate(label = 'test')))
```

Incorporate features discussed in [Exploring Survival on the Titanic](https://www.kaggle.com/mrisdal/titanic/exploring-survival-on-the-titanic/discussion).

Create a Title variable and view its counts by Sex:
```{r}
consol.df %<>%
  dplyr::mutate(Title = gsub('(.*, )|(\\..*)', '', Name))

table(consol.df$Sex, consol.df$Title)
```
Combine rare titles (and reassign some titles) then display the counts again:
```{r}
rare_title <- c(
  'Dona',
  'Lady',
  'the Countess',
  'Capt',
  'Col',
  'Don',
  'Dr',
  'Major',
  'Rev',
  'Sir',
  'Jonkheer'
  )
consol.df %<>%
  dplyr::mutate(Title = ifelse(Title %in% rare_title, 'Rare Title', Title)) %>%
  dplyr::mutate(
  Title = ifelse(Title == 'Mlle', 'Miss', Title),
  Title = ifelse(Title == 'Ms', 'Miss', Title),
  Title = ifelse(Title == 'Mme', 'Mrs', Title)
  )
  
table(consol.df$Sex, consol.df$Title)
```
Create a Surname variable:
```{r}
consol.df %<>%
  dplyr::mutate(Surname = stringr::str_replace(Name, '(.*), .*', '\\1'))

```
There are `r length(unique(consol.df$Surname))` unique usernames.

Create a family size:
```{r}
consol.df %<>%
  dplyr::mutate(Fsize = SibSp + Parch + 1,
                FsizeD = ifelse(Fsize == 1, 'singleton', 'NA'),
                FsizeD = ifelse(Fsize > 1 & Fsize < 5, 'small', FsizeD),
                FsizeD = ifelse(Fsize >= 5, 'large', FsizeD)
  )

```
## Missing Data
View missing data pattern:
```{r}
Amelia::missmap(consol.df)
```
Try to impute Age and Embarked. Give up on Cabin (too many missing values)

### Impute Missing `Embarked` values
View the `Pclass` and `Fare` of the passengers with the missing `Embarked` values:
```{r}
consol.df %>%
  dplyr::filter(is.na(Embarked)) %>%
  dplyr::select(Embarked, Pclass, Fare)
```
Vizualize relationship between `Fare` and `Embarked` accross `Pclass`:
```{r}
ggplot(dplyr::filter(consol.df,!is.na(Embarked)),
       aes(
       x = Embarked,
       y = Fare,
       fill = factor(Pclass)
       )) +
  geom_boxplot() +
  geom_hline(aes(yintercept = 80),
             color = 'red',
             linetype = 'dashed') + 
  scale_y_continuous(labels=dollar_format())
```
Threfore, we assume the `Embarked` value for the two missing values to be *C*:
```{r}
consol.df %<>%
  dplyr::mutate(Embarked = dplyr::case_when(
  is.na(consol.df$Embarked) ~ 'C',
  TRUE ~ as.character(consol.df$Embarked)
  )) %>%
  dplyr::mutate(Embarked = as.factor(Embarked))
  
  summary(consol.df)
```
Fix the one missing value in Fare:
```{r}
consol.df %>%
  dplyr::filter(is.na(Fare)) %>%
  dplyr::select(Fare, Pclass, Embarked)
```
Replace that missing value with the median `Fare` for observations with `Pclass` equals 3 and `Embarked` equals S:
```{r}
mFare <- consol.df %>%
  dplyr::filter(Pclass == '3' & Embarked == 'S' & !is.na(Fare)) %>%
  dplyr::select(Fare) %>%
  {
    median(.$Fare)
  }

consol.df %<>%
  dplyr::mutate(Fare = ifelse(is.na(Fare), mFare, Fare))
```
Create a Child/Not Child variable:
```{r}
consol.df %<>%
  dplyr::mutate(Child = ifelse(Age < 18, 'Child', 'Note Child'))
```
Create a Mother variable:
```{r}
consol.df %<>%
  dplyr::mutate(Mother = ifelse(Sex == 'female' & Parch > 0 & Age > 18 & Title != 'Miss', 'Mother', 'Not Mother'))
```
## Last Steps
Drop unneeded variables:
```{r}
consol.df %<>%
  dplyr::select( -Cabin, -Name, - Ticket, -Surname, -Fsize) 
```
Check variable types are correct:
```{r}
summary(consol.df)
```
```{r}
consol.df %<>%
  dplyr::mutate_each(c('as.factor'), Title:Mother)
```
Split back into the training and test set:
```{r}
train.df <- consol.df %>%
  dplyr::filter(label == 'train') %>%
  dplyr::select(-PassengerId, -label)

test.df <- consol.df %>%
  dplyr::filter(label == 'test') %>%
  dplyr::select(-label)
```
# Prediction
Create an H2O instance:
```{r}
h2o.init(nthreads=-1, max_mem_size='4G')
```
Send `train.df` and `test.df` to h2o:
```{r}
train.hex <- as.h2o(train.df, destination_frame = 'train.hex')
test.hex <- as.h2o(test.df, destination_frame = 'test.hex')
```
Run a deep learning algorithm with random grid search:
```{r}
hyper_params <- list(
  activation = c(
  "Rectifier",
  "Tanh",
  "Maxout",
  "RectifierWithDropout",
  "TanhWithDropout",
  "MaxoutWithDropout"
  ),
  hidden = list(c(200, 200), c(200), c(400), c(400, 200, 100), c(200,200,200)),
  input_dropout_ratio = c(0, 0.05),
  l1 = seq(0, 1e-4, 1e-6),
  l2 = seq(0, 1e-4, 1e-6)
  )
  
  search_criteria <-
  list(
  strategy = "RandomDiscrete",
  max_runtime_secs = 10800,
  max_models = 100,
  seed = 1234567,
  stopping_rounds = 5,
  stopping_tolerance = 1e-2
  )
  
  dl_random_grid <- h2o.grid(
  algorithm = "deeplearning",
  grid_id = "dl_grid_random",
  training_frame = train.hex,
  x = 2:12,
  y = 1,
  epochs = 100,
  stopping_metric = "logloss",
  stopping_tolerance = 1e-2,
  ## stop when logloss does not improve by >=1% for 2 scoring events
  stopping_rounds = 2,
  #score_validation_samples = 10000,
  ## downsample validation set for faster scoring
  score_duty_cycle = 0.025,
  ## don't score more than 2.5% of the wall time
  max_w2 = 10,
  ## can help improve stability for Rectifier
  hyper_params = hyper_params,
  search_criteria = search_criteria
  )
  grid <-
  h2o.getGrid("dl_grid_random", sort_by = "logloss", decreasing = FALSE)
  model.dl <-
  h2o.getModel(grid@model_ids[[1]]) ## model with lowest logloss
  summary(model.dl)

```
Make predictions on the test set:
```{r}
preds <- h2o.predict(model.dl, test.hex)
```
# Submission
Create a file to submit to Kaggle:
  
```{r}
test.df %>%
  dplyr::bind_cols(as.data.frame(preds$predict)) %>%
  dplyr::select(PassengerId, Survived = predict) %>%
  write.csv('kaggle_titanic_submission.csv', row.names = F)
```
